---
title: "Data visualisation"
output: html_notebook
---
## Data visualisation
### Introduction
* Example
```{r}
library(tidyverse)
head(mpg)
ggplot(data=mpg)+
  geom_point(mapping = aes(x=displ,y=hwy))
```
The plot shows a negative relationship between engine size (displ) and fuel efficiency (hwy). 
```{r}
ggplot(data=mpg)
ggplot(data=mpg)+
  geom_point(mapping = aes(x=class,y=drv))
```
### A graphing template
```
ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```
### Aesthetic mappings
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ,y=hwy,color=class))

ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ,y=hwy,shape=class))

ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ,y=hwy,size=class))

ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ,y=hwy,alpha=class))

ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ,y=hwy),color="blue")
```

What happened to the SUVs? ggplot2 will only use six shapes at a time. By default, additional groups will go unplotted when you use the shape aesthetic.

* The name of a color as a character string.
* The size of a point in mm.
* The shape of a point as a number, as shown in [Figure](https://r4ds.had.co.nz/data-visualisation.html#fig:shapes).

```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ,y=hwy,color=displ))

# ggplot(data = mpg)+
#  geom_point(mapping = aes(x=displ,y=hwy,shape=displ))
# Error: A continuous variable can not be mapped to shape

ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ,y=hwy,size=displ))

ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ,y=hwy,color=displ<5))
```
### Common problems
One common problem when creating ggplot2 graphics is to put the + in the wrong place: it has to come at the end of the line, not the start. In other words, make sure you haven’t accidentally written code like this:
```
ggplot(data = mpg) 
+ geom_point(mapping = aes(x = displ, y = hwy))
```

### Facet
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ class)
```
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(. ~ class)
```
If you prefer to not facet in the rows or columns dimension, use a . instead of a variable name, e.g. + facet_grid(. ~ cyl).
### Exercise
1. 
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(.~displ)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~displ)
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))+
  facet_grid(drv~cyl)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)
```

### Geometric objects
```{r}
ggplot(data = mpg)+
  aes(x = displ, y = hwy)+
  geom_point()

ggplot(data = mpg)+
  aes(x = displ, y = hwy)+
  geom_smooth()

ggplot(data = mpg)+
  aes(x = displ, y = hwy,linetype=drv)+
  geom_smooth()

```

```{r}
ggplot(data=mpg)+
  geom_smooth(mapping = aes(x = displ, y = hwy))

ggplot(data=mpg)+
  geom_smooth(mapping = aes(x = displ, y = hwy,group=drv))

ggplot(data=mpg)+
  geom_smooth(mapping = aes(x = displ, y = hwy,color=drv),
              show.legend = F)
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color=class)) + 
  geom_smooth()
```

The local data argument in geom_smooth() overrides the global data argument in ggplot() for that layer only.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color=class)) + 
  geom_smooth(data=filter(mpg,class=="subcompact"),se=F)
```

#### Exercises
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth(se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth(mapping = aes(group=drv),se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy,color=drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color=drv)) + 
  geom_smooth(se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color=drv)) + 
  geom_smooth(mapping = aes(linetype=drv),se = FALSE)

```
### Statistical transformations
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut))
```
Many graphs, like scatterplots, plot the raw values of your dataset. Other graphs, like bar charts, calculate new values to plot:

```{r}
demo <- tribble(
  ~cut,         ~freq,
  "Fair",       1610,
  "Good",       4906,
  "Very Good",  12082,
  "Premium",    13791,
  "Ideal",      21551
)

ggplot(data = demo) +
  geom_bar(mapping = aes(x = cut, y = freq), stat = "identity")
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))
```

```{r}
ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
```

#### Exercises
1. What is the default geom associated with stat_summary()? How could you rewrite the previous plot to use that geom function instead of the stat function?

2. 
```
ggplot(data = diamonds) + 
  geom_col(mapping = aes(x=cut))
```
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop..,group=1))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop..,group=1))
```

### Position adjustments
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity))
```
```{r}
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) + 
  geom_bar(alpha = 1/5, position = "identity")
ggplot(data = diamonds, mapping = aes(x = cut, colour = clarity)) + 
  geom_bar(fill = NA, position = "identity")

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), position = "jitter")
```

#### Exercises
```{r}
ggplot(data = mpg, mapping = aes(x = cty)) + 
  geom_bar()
```
### Coordinate systems
```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot()

ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot()+
  coord_flip()
```

```{r}
nz <- map_data("nz")
ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "blue")

ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black") +
  coord_quickmap()
```

```{r}
bar <- ggplot(data = diamonds) + 
  geom_bar(
    mapping = aes(x = cut, fill = cut), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)
bar
bar + coord_flip()
bar + coord_polar()
```
#### Exercises
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity))+
  labs(x=NULL,y=NULL)+
  theme(aspect.ratio = 1) +
  coord_polar()
```

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline()+
  coord_fixed()
```
fixed the scale.
### The layered grammar of graphics
```
ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(
     mapping = aes(<MAPPINGS>),
     stat = <STAT>, 
     position = <POSITION>
  ) +
  <COORDINATE_FUNCTION> +
  <FACET_FUNCTION>
```
### References
1. [Data visualisation](https://r4ds.had.co.nz/data-visualisation.htm)





