---
title: "Graphics for communication"
output: html_notebook
---
## Scales
```{r}
library(tidyverse)
ggplot(mpg,aes(displ, hwy))+
  geom_point(aes(colour=class))

library(tidyverse)
ggplot(mpg,aes(displ, hwy))+
  geom_point(aes(colour=class))+
  scale_y_continuous(breaks = seq(15,50,by=5))

library(tidyverse)
ggplot(mpg,aes(displ, hwy))+
  geom_point(aes(colour=class))+
  scale_x_continuous(labels = NULL)+
  scale_y_continuous(labels = NULL)
```
Collectively axes and legends are called guides. Axes are used for x and y aesthetics; legends are used for everything else.

```{r}
presidential %>%
  mutate(id = 33 + row_number()) %>%
  ggplot(aes(start, id)) +
    geom_point() +
    geom_segment(aes(xend = end, yend = id)) +
    scale_x_date(NULL, breaks = presidential$start, date_labels = "'%y")
```

```{r}
base <- ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class))

base + theme(legend.position = "left")
base + theme(legend.position = "top")
base + theme(legend.position = "bottom")
base + theme(legend.position = "right") # the default
```

```{r}
ggplot(mpg,aes(displ,hwy))+
  geom_point(aes(color=class))+
  geom_smooth(se=F)+
  theme(legend.position = "bottom")+
  guides(color=guide_legend(nrow=1,override.aes = list(size = 4)))
```

```{r}
ggplot(diamonds, aes(log10(carat), log10(price))) +
  geom_point()

ggplot(diamonds, aes(log10(carat), log10(price))) +
  geom_bin2d()
```

```{r}
ggplot(diamonds, aes(carat, price)) +
  geom_bin2d() + 
  scale_x_log10()+
  scale_y_log10()
```

* Color
```{r}
ggplot(mpg,aes(displ,hwy))+
  geom_point(aes(color=drv))

ggplot(mpg,aes(displ,hwy))+
  geom_point(aes(color=drv))+
  scale_color_brewer(palette = "Set1")

ggplot(mpg,aes(displ,hwy))+
  geom_point(aes(color=drv))+
  scale_color_brewer(palette = "Set2")

ggplot(mpg,aes(displ,hwy))+
  geom_point(aes(color=drv,shape=drv))+
  scale_color_brewer(palette = "Set2")
```

```{r}
presidential%>%
  mutate(id=33+row_number())%>%
  ggplot(aes(start,id,color=party))+
  geom_point()+
  geom_segment(aes(xend=end,yend=id))+
  scale_color_manual(values = c(Republican = "red", Democratic = "blue"))
```

```{r}
df <- tibble(
  x=rnorm(10000),
  y=rnorm(10000)
)

ggplot(df,aes(x,y))+
  geom_hex()+
  coord_fixed()

ggplot(df, aes(x, y)) +
  geom_hex() +
  viridis::scale_fill_viridis() +
  coord_fixed()

ggplot(df, aes(x, y)) +
  geom_hex() +
  scale_color_gradient(low = "white",high = "red")+
  coord_fixed()
```

### Exercises
1. Why doesn’t the following code override the default scale?
```{r}
df <- tibble(
  x=rnorm(10000),
  y=rnorm(10000)
)

ggplot(df, aes(x, y)) +
  geom_point(aes(color=x))+
  scale_colour_gradient(low = "white", high = "red") +
  coord_fixed()

ggplot(df, aes(x, y)) +
  geom_hex()+
  scale_colour_gradient(low = "white", high = "red") +
  coord_fixed()

```

Note that all colour scales come in two variety: scale_colour_x() and scale_fill_x() for the colour and fill aesthetics respectively (the colour scales are available in both UK and US spellings).

4. Use override.aes to make the legend on the following plot easier to see.
ggplot(diamonds, aes(carat, price)) +
  geom_point(aes(colour = cut), alpha = 1/20)
```{r}
ggplot(diamonds, aes(carat, price)) +
  geom_point(aes(colour = cut), alpha = 1/20)+
  guides(colour = guide_legend( override.aes = list(size = 4,alpha=1)))
```
### Zomming
```{r}
ggplot(mpg,mapping = aes(displ,hwy))+
  geom_point(aes(color=class))+
  geom_smooth()+
  coord_cartesian(xlim=c(5,7),ylim = c(10,30))

mpg%>%
  filter(displ>=5,displ<=7,hwy>=10,hwy<=30)%>%
  ggplot(aes(displ,hwy))+
  geom_point(aes(color=class))+
  geom_smooth()
```


```{r}
suv <- mpg %>% filter(class == "suv")
compact <- mpg %>% filter(class == "compact")

ggplot(suv, aes(displ, hwy, colour = drv)) +
  geom_point()

ggplot(compact, aes(displ, hwy, colour = drv)) +
  geom_point()
```
```{r}
x_scale <- scale_x_continuous(limits = range(mpg$displ))
y_scale <- scale_y_continuous(limits = range(mpg$hwy))
col_scale <- scale_colour_discrete(limits = unique(mpg$drv))

ggplot(suv, aes(displ, hwy, colour = drv)) +
  geom_point() +
  x_scale +
  y_scale +
  col_scale

ggplot(compact, aes(displ, hwy, colour = drv)) +
  geom_point() +
  x_scale +
  y_scale +
  col_scale
ggsave("my-plot.pdf")
```

#### Figure sizing

### References
1. [graphics-for-communication](https://r4ds.had.co.nz/graphics-for-communication.html)



